---
title: "Global CO2 Emissions Data Analysis"
author: "Jennifer Liu, David Kook, Xinwen Xu, Yuxuan Peng, William Escobar"
date: "2022-11-10"
output: pdf_document
---

# Total CO2 Emission Analysis using data from 1750-2020
The code below analyzes CO2 emission data from Kaggle. The three main goals are:
1. to understand the historical cumulative emissions, \
2. Present day emission analysis,\
3. Using this data for future emission predictions. 

## Libraries needed for this analysis:
```{r setup, include=FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plyr)
library(dplyr)
library(data.table)
library(countrycode)
library(ggrepel)
library(magick)
library(scales)
library(ggplot2)
library(ggpubr)
library(maptools)
library(rworldmap)
library(RCurl)
library(httr)
library(sqldf)
library(forecast)
library(tmap)
library(sf)
library(raster)
library(spData)
library(devtools)
library(RColorBrewer)
library(magrittr)
library(heatmaply)
```

## Data Cleaning
```{r Read in the raw data}
data <- getURL("https://raw.githubusercontent.com/jliuob/QBSsquad/main/CO2_emission_by_countries.csv")
rawDat <- read.csv(text = data)
```

```{r Check for all the NA entries}
rawCO2<-rawDat 
rawCO2%>% summarise_all(~ sum(is.na(.)))
```

```{r NA Countries Within Columns}
#Find all the countries that has NA in each column
naCode <- rawCO2 %>% filter(is.na(Code)) %>% distinct(Country) 
naCall <- rawCO2 %>% filter(is.na(Calling.Code)) %>% distinct(Country)
naPopu <- rawCO2 %>% filter(is.na(Population.2022.)) %>% distinct(Country)
naArea <- rawCO2 %>% filter(is.na(Area)) %>% distinct(Country)
naXPre <- rawCO2 %>% filter(is.na(X..of.World)) %>% distinct(Country)
naDens <- rawCO2 %>% filter(is.na(Density.km2.)) %>% distinct(Country)
allNA <- bind_rows(naCode,naCall,naPopu,naArea,naXPre,naDens)
allNA <- distinct(allNA) %>% arrange(Country)
head(allNA,10) #In total of 41 countries have different missing values
```

Combine all the table above to better see which values are missing.
```{r Missing Value table}
nacheck<-function(str){ 
  #This function get the individual column data above, 
  df<- get(str)
  df$UniC <- df$Country
  #create a second column to later join all the table together
  colnames(df)[1] = str 
  #rename the country column as the abbreviation of the initial data frame
  return(df)
}
#run the function for all the column that had missing data
naCode<- nacheck("naCode")
naCall<- nacheck("naCall")
naPopu<- nacheck("naPopu")
naArea<- nacheck("naArea")
naXPre<- nacheck("naXPre")
naDens<- nacheck("naDens")
#Join all the column together
allNATable<-allNA %>% full_join(naCode,by = c("Country" = "UniC")) %>% 
  full_join(naCall,by = c("Country" = "UniC")) %>%
  full_join(naPopu,by = c("Country" = "UniC")) %>%
  full_join(naArea,by = c("Country" = "UniC")) %>%
  full_join(naXPre,by = c("Country" = "UniC")) %>%
  full_join(naDens,by = c("Country" = "UniC"))
view(allNATable)
```

```{r}
write.csv(allNATable, file = "OriginalNAs.csv",row.names = TRUE)
#Write out the table for future reference, can also see this data in Excel
```

```{r Find the threshold}
dat20 <- rawCO2 %>% filter(Year == 2020)
summary(dat20$CO2.emission..Tons.)
#decided on the first quartile 
```

```{r Finalize Cleaning}
#Read in the cleaned version from excel
y <- getURL("https://raw.githubusercontent.com/jliuob/QBSsquad/main/CO2_Cleaned_V2.csv")
CO2 <- read.csv(text = y)
colnames(CO2)[1] = "Country"
#again, check for all the NA entries
CO2 %>% summarise_all(~ sum(is.na(.)))
CO2 <- CO2 %>% dplyr::select(-c(8,9))
#omitted the last two column 
head(CO2)
#check the dimension of data
dim(CO2)
```
### Please proceed with this CO2 (don't forget to omit.na)


## Historical CO2 Data analysis and visualization
## Aim 1.1 General overview of data, any visual or numerical differences.
Finding the first year in history that countries began emitting CO2
```{r General Overview}
#finding and removing NAs
CO2 <- na.omit(CO2)
View(CO2)

#filter emissions 0 and use tidyverse groupby country minimum year
CO2_no0<-CO2 %>% group_by(min(Year)) %>% filter(CO2.emission..Tons.!=0)

CO2_no0 %>% 
    group_by(Country) %>% 
    slice(which.min(Year))

CO2_no0 <- data.table(CO2_no0)

#removing duplicates country names
CO2_noduplicates <- CO2_no0[!duplicated(CO2_no0$Country),]
CO2_noduplicates 

CO2_no_year_min <- CO2_no0[, list(Year = min(Year)), by = Country] 
#new data table to identify first year with non-zero CO2 emission value
CO2_no_year_min[]
summary(CO2_no_year_min$Year)

#scatter plot of the years with non-zero CO2 value for different countries, with line at 1950
plot(CO2_no_year_min$Year,main="Earliest Year with CO2 Data",xlab="Index",ylab="Year")+abline(h=1950,col="blue")

```
Based on the scatter plot above, in our data, most countries have their first non-zero CO2 value at 1950. 

### Finding total CO2 emissions per country & CO2 emissions per country per person
```{r total emissions per country}
#finding CO2 emission/population
CO2total_byCountry <- aggregate(CO2_no0$CO2.emission..Tons., by=list(Category=CO2_no0$Country), FUN=sum) 

#renaming columns
colnames(CO2total_byCountry)[1] <- "Country"
colnames(CO2total_byCountry)[2] <- "CO2.Total"
CO2total_byCountry #checking that the column renaming worked

#joining columns together to conduct calculations
CO2total_byCountry <- CO2total_byCountry %>% inner_join(CO2_noduplicates) %>% select("Country","CO2.Total","Population.2022.")

#dividing total CO2 emissions per country by the country's population
CO2total_byCountry$CO2perPop <- NULL
CO2total_byCountry$CO2perPop <- CO2total_byCountry$CO2.Total/CO2total_byCountry$Population.2022.
CO2total_byCountry

#renaming column names
colnames(CO2total_byCountry)[2] <- "CO2.Total.Per.Country"
colnames(CO2total_byCountry)[4] <- "CO2.per.Person.per.Country"

summary(CO2total_byCountry)
CO2total_byCountry 

```
The results of `summary()` on the CO2total_byCountry data set indicate that the average total CO2 emissions in the data set is 3.024e+11 tons. 
 
## Aim 1.2 A: To visualize the cumulative CO2 emissions by country and continent
```{r,message=FALSE}
CO2total_byCountry$continent <- countrycode(sourcevar = CO2total_byCountry[,"Country"],
                                  origin = "country.name",
                                  destination = "continent")

#summing the total CO2 emissions by geographic region: Oceania, Africa, Asia, Americas, and Europe (Antartica was not included in the dataset)
CO2total_byContinent <- aggregate(CO2total_byCountry$CO2.Total.Per.Country, by=list(Category=CO2total_byCountry$continent), FUN=sum)
CO2total_byContinent 
colnames(CO2total_byContinent) <- c("Geographic_Region","Total_CO2_Emissions")

CO2total_byContinent <-as.data.frame(CO2total_byContinent)

continent_barplot <- ggbarplot(data=CO2total_byContinent,x="Geographic_Region",y="Total_CO2_Emissions",fill="Geographic_Region")
continent_barplot <- continent_barplot + labs(x="Continent",
y="Total CO2 Emissions (Tons)",title="CO2 Totals by Continent",fill="Continent")
continent_barplot


```

### Create a map visualizing the amount of CO2 emission by country
To visualize the CO2 emissions by country using a world map, load in ggplot2's map data via the code below. This data set contains longitude, latitude, and country codes to plot the world map. The cleaned CO2 data set will merge with the world data set by country name. 
```{r,message=FALSE}
world <- map_data("world")
#table(world) #world data codes country names differently. Change prior to joining 
#Add the data you want to map countries by to world
#In this example, I add lengths of country names plus some offset
```

### Creating a barplot & world map to display CO2 emissions per country 
The world data set codes some country names differently. So, prior to joining the CO2 and world datasets, match the country names for "United States" and "Democratic Republic of the Congo" to the way the world dataset codes these countries. 
```{r,message=FALSE}
which(CO2total_byCountry$Country=="United States")
which(CO2total_byCountry$Country=="Democratic Republic of Congo")
CO2total_byCountry$Country[196] <- "USA" #renaming "United States" to "USA", to ensure uniformity in the naming 
CO2total_byCountry$Country[47] <- "Democratic Republic of the Congo"
CO2total_byCountry[196,]

#renaming "region" in the data we are importing to "country"
world <- inner_join(world,CO2total_byCountry,by=c("region"="Country"))
#remove duplicates
world <- world[!duplicated(world),]
world$CO2.Total.Per.Country[world$region=="USA"] 

#barplot of CO2 total emissions in world map
continent_barplot <- ggplot(data=subset(CO2total_byCountry,!is.na(continent)),aes(x=continent,y=CO2.Total.Per.Country,fill=continent)) + geom_bar(stat="identity",position="dodge")

continent_barplot <- continent_barplot + labs(x="Continent",y="Total CO2 Emissions (Tons)",title="CO2 Totals by Continent",fill="Continent")
continent_barplot

#graphically presenting CO2 emission in a world map
gg <- ggplot() + theme(legend.position="bottom",legend.key.size = unit(0.5, "cm"),legend.key.width = unit(1,"cm")) + labs(fill="CO2 Total(tons)/Country",x="Longitude",y="Latitude")
gg <- gg + geom_map(data=world, map=world, aes(map_id=region,x=long, y=lat, fill=CO2.Total.Per.Country))

#changing features of world map (color, axes scales, adding a title, etc)
gg <- gg + scale_fill_gradient(low = "green", high = "red", guide = "colourbar",aesthetics = "fill") + scale_y_continuous(limits = c(-100,100)) + scale_x_continuous(limits = c(-175,175)) + scale_colour_continuous(labels="scientific")
gg <- gg + coord_equal() + ggtitle("World View of CO2 Emissions")
gg

```

### Making a world map just for 2019 data:

To see if there's a difference between historical cumulative CO2 and 2019
CO2 emissions, we made a world map just for 2019 emissions data. 
```{r,message=FALSE}
cleanedC02 <- CO2 
cleanedC02_2019<-cleanedC02[cleanedC02$Year == "2019",] #below code is same as cumulative analysis
CO2total_byCountry_2019 <- aggregate(cleanedC02_2019$CO2.emission..Tons., by=list(Category=cleanedC02_2019$Country),FUN=sum)
colnames(CO2total_byCountry_2019)[1] <- "Country"
colnames(CO2total_byCountry_2019)[2] <- "CO2.Total"
CO2total_byCountry_2019<- CO2total_byCountry_2019 %>%
  inner_join(cleanedC02_2019) %>%
  select("Country","CO2.Total","Population.2022.")
which(CO2total_byCountry_2019$Country=="United States")
which(CO2total_byCountry_2019$Country=="Democratic Republic of Congo")

CO2total_byCountry_2019$Country[196] <- "USA"#coding USA differently #because world data has it as USA, but CO2 data has it as United States
CO2total_byCountry_2019$Country[47] <- "Democratic Republic of the Congo"
world2 <- map_data("world")
world2 <- inner_join(world2,CO2total_byCountry_2019,by=c("region"="Country"))

gg2 <- ggplot() + theme(legend.position="bottom",legend.key.size = unit(0.5, "cm"),legend.key.width = unit(1,"cm")) + labs(fill="CO2 Total(tons)/Country",x="Longitude",y="Latitude")
gg2 <- gg2 + geom_map(data=world2, map=world2, aes(map_id=region,x=long, y=lat, fill=CO2.Total))

#changing features of world map (color, axes scales, adding a title, etc)
gg2 <- gg2 + scale_fill_gradient(low = "green", high = "red", guide = "colourbar",aesthetics = "fill") + scale_y_continuous(limits = c(-100,100)) + scale_x_continuous(limits = c(-175,175)) + scale_colour_continuous(labels="scientific")
gg2 <- gg2 + coord_equal() + ggtitle("World View of CO2 Emissions in 2019")
gg2


```

### Analyze the top 3 countries that emit the most CO2 by geographic region
```{r,message=FALSE}
#finding and plotting the top 3 countries in each geographic region that emit the most CO2
topCountries <- CO2total_byCountry[order(-CO2total_byCountry$CO2.Total.Per.Country),] #get CO2 emissions per country in descending order

#finding the top 3 countries in each geographic region that emit the most CO2
top3Africa<-head(topCountries[topCountries$continent=="Africa",],3)
top3Americas<-head(topCountries[topCountries$continent=="Americas",],3)
top3Asia<-head(topCountries[topCountries$continent=="Asia",],3)
top3Europe<-head(topCountries[topCountries$continent=="Europe",],3)
top3Oceania<-head(topCountries[topCountries$continent=="Oceania",],3)

top3pContinent <- rbind(top3Africa,top3Americas,top3Asia,top3Europe,top3Oceania)

#soring the data into a variable for plotting
mostEmissions <- ggplot(top3pContinent,aes(x=continent,y=CO2.Total.Per.Country))

#plotting the data
mostEmissions<- mostEmissions +
  geom_text_repel(aes(label = Country), size = 2.5)+
  geom_point(aes(colour=CO2.Total.Per.Country))+
  ggtitle(expression(atop("Top 3 CO2 Country Emissions per Continent",
                          atop(italic("CO2 Total from Available Data"),""))))+
  xlab("Continent") +
  ylab("CO2 Emissions in Tons")+
  theme(text = element_text(family = 'serif'))+
  theme(plot.title = element_text(hjust=0.5))

mostEmissions

```

## 1.2B: C02 Emissions by Area:
```{r CO2 Emissions by Area,message=FALSE}
cleanedC02 <- CO2 
#identified countries with NA values, then excluded the lower 50% of the countries who contributed the least CO2 emissions


#joining country, total area of each country, percent of global land mass that the country takes up, and the total CO2 emissions per country into one table


cleanedC02 <- inner_join(cleanedC02,CO2total_byCountry) %>% select("Country","Area","CO2.Total.Per.Country","Population.2022.")


#plotting a scatter plot for country land mass vs. total CO2 emissions per, and finding the correlation coefficient (R^2 value), to assess the strength of the correlation
areavsco2plot <- ggplot(data=cleanedC02,aes(x=Area,y=CO2.Total.Per.Country))+
  geom_point(aes(fill=CO2.Total.Per.Country)) +
  ggtitle("Country Area vs. Total CO2 Emission") +
  labs(x="Area",y="Total CO2 Emissions (Tons)",
       title="CO2 Emissions vs. Country Area (km2)") + 
  geom_text(aes(label=ifelse(Area>5e6,as.character(Country),'')),size=2.5,hjust=1,vjust=-1) +
  stat_cor(method = "pearson",label.x = 0e0, label.y = 2e13)

areavsco2plot
```

### C02 Emissions in Developed vs. Non-developed:
```{r}
x <- getURL("https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv")
dat2 <- read.csv(text = x)
y <- getURL("https://raw.githubusercontent.com/jliuob/QBSsquad/main/CO2_Cleaned_V2.csv")
dat1 <- read.csv(text = y)
dat1 <- dat1[,1:7]
class(dat1)
class(dat2)
colnames(dat1)[1] <- "Country"
```

```{r}
merged.dat <- sqldf("SELECT d1.Country, d1.year, d1.'CO2.emission..Tons.',
d2.gdp, d2.co2_per_gdp, d2.co2_per_capita, d2.coal_co2, d2.coal_co2_per_capita
      FROM 'dat1' AS d1
      JOIN 'dat2' AS d2
      ON d1.Country=d2.country
      AND d1.Year=d2.year")
dim(merged.dat)
merged.dat <- na.omit(merged.dat)
View(merged.dat)
```

### Finding CO2 per capita per country:
```{r}
#finding the average CO2 emissions per country
avgCO2byCountry <- aggregate(merged.dat$CO2.emission..Tons.,by=list(Category=merged.dat$Country), FUN=mean)

#finding the average gdp per country
#recode NAs to zero to run mean functions
merged.dat$gdp[is.na(merged.dat$gdp)] <- 0
avgGDP <- aggregate(merged.dat$gdp,by=list(Category=merged.dat$Country), FUN=mean)

merged.dat$gdp[merged.dat$Country=="Aruba"]

#calculating the gdp per capita for each country
avgC02perCapita <- aggregate(merged.dat$co2_per_capita,by=list(Category=merged.dat$Country), FUN=mean)

#omitting na values and finding the top 10 CO2 emissions per capita per country 
avgC02perCapita <-na.omit(avgC02perCapita)
avgC02perCapita_10 <- avgC02perCapita[order(avgC02perCapita$x),]
avgC02perCapita_10 <- head(avgC02perCapita_10,10)

#plotting top 10 CO2 emissions per capita per country
capitaplot <- ggplot(avgC02perCapita_10,
                     aes(Category,x,fill=Category))+
  geom_bar(stat="identity")+
  scale_x_discrete(guide=guide_axis(angle = 90))+
  xlab("Country")+
  ylab("C02 per Capita")+
  ggtitle(expression(atop("C02 per Capita per Country",
                          atop(italic("Top 10 Countries"),"")))) +
  theme(plot.title = element_text(hjust=0.5))+
  theme(text = element_text(family = 'serif'))
capitaplot
```

### C02 by GDP:
```{r}
colnames(avgGDP) <- c("Country","GDP_Average") #renaming column names

#merging country, GDP_Average, and Population 2022 columns into one table
avgGDP <- avgGDP %>% inner_join(cleanedC02) %>% select("Country","GDP_Average","Population.2022.")
avgGDP <- avgGDP %>% mutate(GDPperCapita = GDP_Average/Population.2022.) 

#creating ordinal data for developed vs. developing
#changed developed threshold to $9000 per capita because original GDP threshold was too conservative
develop_ord <- ifelse(avgGDP$GDPperCapita>=9000,"developed",
                      ifelse(avgGDP$GDPperCapita<9000,"developing",0))
avgGDP$DevelopStatus <- develop_ord

table(avgGDP$Country[avgGDP$GDPperCapita>=9000])
table(avgGDP$Country[avgGDP$GDPperCapita<9000])

avgGDP$GDP_Average

#merging C02 Data
avgGDP <- avgGDP %>% inner_join(CO2_noduplicates) %>% select("Country","GDP_Average","Population.2022.","GDPperCapita","CO2.emission..Tons.","DevelopStatus")

#categorizing developed nations into a developed variable
#categorizing developing nations into a developing variable
developed <- avgGDP[avgGDP$DevelopStatus=="developed",]
developing <- avgGDP[avgGDP$DevelopStatus=="developing",]
sumC02Developed <- sum(developed$CO2.emission..Tons.)
sumC02Developing <- sum(developing$CO2.emission..Tons.)

#creating a table and a barplot comparing developed vs developing nations total CO2 emissions
DevelopvsDeveloping <- matrix(ncol=2,nrow=2)
DevelopvsDeveloping <- as.data.frame(DevelopvsDeveloping)
colnames(DevelopvsDeveloping)<-c("CO2_Total","Status")
DevelopvsDeveloping[1,] <- c(as.numeric(sumC02Developed),"Developed")
DevelopvsDeveloping[2,] <- c(as.numeric(sumC02Developing),"Developing")

DevelopvDevelopingplot <- ggplot(DevelopvsDeveloping,
                                 aes(x=Status,y=CO2_Total,fill=Status))+
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(size = 11)) +
  scale_x_discrete(guide = guide_axis(n.dodge = 3))+
  xlab(c("Development Status"))+
  ylab("C02 Total Tons")+
  ggtitle("C02 Emission Differences by Development Status")+
  theme_gray()+
  theme(plot.title = element_text(hjust=0.5))+
  theme(text = element_text(family = 'serif'))+
  stat_cor(method = "pearson", label.x = 0, label.y = 0)

DevelopvDevelopingplot
```
There are more countries.

## Aim 2.1 COVID CO2 emission analysis
First, to examine global CO2 emission in pre-pandemic and pandemic years, a subset of data about 2018, 2019, and 2020 CO2 emission in all countries is created.
```{r subset pandemic years}
pandemic.CO2 <- CO2 %>%
  filter(Year >= 2018, na.rm = TRUE)
head(pandemic.CO2)
```

Then, we want to explore the general trend of changes in CO2 emission in all countries due to the pandemic. Thus, we presented the global cumulative CO2 emission during these three years.
```{r cummulative co2 sum}
cum.2018 <- sum(pandemic.CO2[which(pandemic.CO2$Year=='2018'), 5])
cum.2019 <- sum(pandemic.CO2[which(pandemic.CO2$Year=='2019'), 5])
cum.2020 <- sum(pandemic.CO2[which(pandemic.CO2$Year=='2020'), 5])

barplot(tapply(pandemic.CO2$CO2.emission..Tons., format(pandemic.CO2$Year), FUN=sum), col="#69b3a2")
```

We could see from the graph that there's not much change in global cumulative CO2 emission during the pandemic years. Here we want to separately analyze the top5 emission countries and see if there's any difference.
```{r cummlative co2 in top5}
# dataframe for top 5 co2 emission countries each year
CO2.a <- pandemic.CO2 %>% 
  filter(Year == "2018") %>%
  arrange(desc(CO2.emission..Tons.)) %>%
  slice_head(n = 5)
CO2.a

CO2.b <- pandemic.CO2 %>% 
  filter(Year == "2019") %>%
  arrange(desc(CO2.emission..Tons.)) %>%
  slice_head(n = 5)
CO2.b

CO2.c <- pandemic.CO2 %>% 
  filter(Year == "2020") %>%
  arrange(desc(CO2.emission..Tons.)) %>%
  slice_head(n = 5)
CO2.c
df.top5 = rbind(CO2.a,CO2.b,CO2.c)

# plot general trend
ggplot(df.top5, aes(x = Year, y = CO2.emission..Tons., color = Country)) +
  geom_line()
```
As shown in this graph, although the overall CO2 emission in these 5 top emission countries slightly increased during pandemic, there's still not much difference observed. This is not what we've expected before since we expect the emission rate to decrease during the pandemic.\
\
Now we want to join another dataset with information about absolute growth of CO2 and other information about CO2 per capita/gdp/unit energy to explore further on pandemic-CO2 emission relationship.\
\
First, we read our new dataset about owid-co2-data from github and filtered the wanted period.
```{r new data set}
res = GET("https://github.com/owid/co2-data")

x <- getURL("https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv")
world.co2 <- read.csv(text = x)
class(world.co2)
world.pandemic <- world.co2 %>%
  filter(year >= 1990, na.rm = TRUE)
```

Then we joined the two datasets based on country names. Here we used SQL to make the join.
```{r sql join}
# based on both country names and year
tot.dat <- sqldf("SELECT p.Country, p.year, p.'CO2.emission..Tons.' AS 'co2 emission(tons)', p.'Population.2022.' AS population, p.Area, w.co2_growth_abs, w.co2_growth_prct, w.co2_per_capita
      FROM 'pandemic.CO2' AS p
      JOIN 'world.pandemic' AS w 
      ON p.Country=w.country
      AND p.Year=w.year")
```

Now we wanted to plot the absolute growth of co2 emission during the pandemic in a global perspective. First, let's check if there are any missing values in our interested data.
```{r check NA}
sum(is.na(tot.dat$co2_growth_abs))
which(is.na(tot.dat$co2_growth_abs))

# NAs are located at Antarctica, Christmas Island, and Puerto Rico. We noticed that there might be no data available for those countries so we just denoted them as 0.
tot.dat$co2_growth_abs[is.na(tot.dat$co2_growth_abs)] = 0

#check again
sum(is.na(tot.dat$co2_growth_abs))
```

Now we can visualize our results regarding absolute growth of CO2 emission. First, let's check if there's difference in the absolute CO2 growth during the pandemic.\
\
Then we plotted the absolute co2 growth and co2 emission per-capita during the pandemic in countries with the highest overall co2 emission these years.
```{r abs co2 visualization}
# global trend in co2 absolute growth
a <- sum(tot.dat[which(tot.dat$Year=='2018'), 6])
b <- sum(tot.dat[which(tot.dat$Year=='2019'), 6])
c <- sum(tot.dat[which(tot.dat$Year=='2020'), 6])
co2.gr <- c(a,b,c)
plot(co2.gr, type = "line")

# trend in top 10 co2 emission countries
tot.dat <- tot.dat %>%
  arrange(desc(`co2 emission(tons)`))
top.co2 <- tot.dat[1:30,]

ggplot(top.co2, aes(x = Year, y = co2_growth_abs, color = Country)) +
   xlab("Year")+
  ylab("CO2 Absolute Growth")+
  ggtitle("Absolute CO2 Growth During Pandemic Years in top10 CO2 Emission Countries")+
  geom_line()
ggplot(top.co2, aes(x = Year, y = co2_per_capita, color = Country)) +
   xlab("Year")+
  ylab("CO2 per capita")+
  ggtitle("CO2 Emission per capita During Pandemic Years in top10 CO2 Emission Countries")+
  geom_line()
```

To further explore the relationship and intensity of different features in our dataset, we filtered out top 10 co2 emission countries each year and presented our observations in a heatmap.
```{r heatmap top10 emission}
# filtered out top 10 emission countries
CO2.1 <- tot.dat %>% 
  filter(Year == "2018") %>%
  arrange(desc(`co2 emission(tons)`)) %>%
  slice_head(n = 10)


CO2.2 <- tot.dat %>% 
  filter(Year == "2019") %>%
  arrange(desc(`co2 emission(tons)`)) %>%
  slice_head(n = 10)


CO2.3 <- tot.dat %>% 
  filter(Year == "2020") %>%
  arrange(desc(`co2 emission(tons)`)) %>%
  slice_head(n = 10)

co2.top10 = rbind(CO2.1,CO2.2,CO2.3)
co2.top10$Country_Year <- paste(co2.top10$Country, co2.top10$Year)
rownames(co2.top10) <- co2.top10[, 9]

# each country along with the year should be distinctly represented
co2.top10 %<>% arrange(Country_Year)

# interactive heatmap
dir.create("folder")
heatmaply(percentize(co2.top10)[-(1:2)],
          xlab = "Features",
          ylab = "Country Year",
          main = "Features related to CO2 emission during pandemic in top10 CO2 emission countries",
          margins = c(60,100,40,20),
          colors = colorRampPalette(brewer.pal(3, "RdBu"))(256),
          seriate = "OLO",
          heatmap_layers = theme(axis.line=element_blank()),
          file = "folder/CO2_plot.html")
browseURL("folder/CO2_plot.html")
```

We did the same thing by filtered out top 10 populated countries each year and presented our observations in a heatmap.
```{r heatmap top10 populated}
# filtered out top 10 populated countries
pop1 <- tot.dat %>% 
  filter(Year == "2018") %>%
  arrange(desc(population)) %>%
  slice_head(n = 10)


pop2 <- tot.dat %>% 
  filter(Year == "2019") %>%
  arrange(desc(population)) %>%
  slice_head(n = 10)


pop3 <- tot.dat %>% 
  filter(Year == "2020") %>%
  arrange(desc(population)) %>%
  slice_head(n = 10)

pop.top10 = rbind(pop1,pop2,pop3)
pop.top10$Country_Year <- paste(pop.top10$Country, pop.top10$Year)
rownames(pop.top10) <- pop.top10[, 9]

# each country along with the year should be distinctly represented
pop.top10 %<>% arrange(Country_Year)

# interactive heatmap
dir.create("folder")
heatmaply(percentize(pop.top10)[-(1:2)],
          xlab = "Features",
          ylab = "Country Year",
          main = "Features related to CO2 emission during pandemic in top10 populated countries",
          margins = c(60,100,40,20),
          colors = colorRampPalette(brewer.pal(3, "RdBu"))(256),
          seriate = "OLO",
          heatmap_layers = theme(axis.line=element_blank()),
          file = "folder/pop_plot.html")
browseURL("folder/pop_plot.html")
```
From the analysis of co2 emissions during COVID, although we cannot conclude an association between changes in co2 emissions and the pandemic because we ignored other variables, we thought it is interesting to display the result. The result might be more significant if we have updated data including 2021 and 2022.


## Aim 2.2 Forecasting 
We decided to predict the CO2 emissions in the next 30 years by using the data from 1992 to 2019, and fit in the proper ARIMA models to do the prediction. We choose six countries and five continents to do the prediction.

```{r extract and prepare the data for forecasting}
# select three columns to do forecasting
CO2_forecasting<-CO2 %>% dplyr::select(Country, Year, CO2.emission..Tons.)
# make sure there is no NA
dim(CO2_forecasting)
dim(na.omit(CO2_forecasting))
# see which year should we start for forecasting
CO2_no_year_min[which(CO2_no_year_min$Year==max(CO2_no_year_min[]$Year)),]
# decided to use 1992-2019 data to do the forecasting
CO2_forecasting<-CO2_forecasting %>% 
  filter(Year>1991&Year<2020) %>% 
  group_by(Country)
# View(CO2_forecasting)
```

### Predictions on selected countries
#### Time series analysis based on six selected countries
```{r forecast on 6 countries}
# extract the six countries' data and fit in time series models separately
# US
CO2_US<-CO2_forecasting %>%
  filter(Country=="United States")
pred.us<-ts(CO2_US$CO2.emission..Tons., start=1992,frequency = 1) %>% 
  auto.arima() %>% 
  forecast(h=30)
# United Kingdom
CO2_UK<-CO2_forecasting %>%
  filter(Country=="United Kingdom")
pred.uk<-ts(CO2_UK$CO2.emission..Tons.,start=1992,frequency = 1) %>% 
  auto.arima() %>% 
  forecast(h=30)
# Germany
CO2_GE<-CO2_forecasting %>%
  filter(Country=="Germany")
pred.ge<-ts(CO2_GE$CO2.emission..Tons.,start=1992,frequency = 1) %>% 
  auto.arima() %>% 
  forecast(h=30)
# China
CO2_CH<-CO2_forecasting %>%
  filter(Country=="China")
pred.ch<-ts(CO2_CH$CO2.emission..Tons.,start=1992,frequency = 1) %>% 
  auto.arima() %>% 
  forecast(h=30)
# Russia
CO2_RU<-CO2_forecasting %>%
  filter(Country=="Russia")
pred.ru<-ts(CO2_RU$CO2.emission..Tons.,start=1992,frequency = 1) %>% 
  auto.arima() %>% 
  forecast(h=30)
# France
CO2_FR<-CO2_forecasting %>%
  filter(Country=="France")
pred.fr<-ts(CO2_FR$CO2.emission..Tons.,start=1992,frequency = 1) %>% 
  auto.arima() %>% 
  forecast(h=30)
```

#### Plot the predicted CO2 emissions for the six countries 
We expect to see what's the increasing or decreasing trend of the CO2 emissions in the next 30 years. 
```{r plot the predictions of the six countries}
# devtools::install_github('cttobin/ggthemr')
library(ggthemr)
ggthemr("dust")
US.plot<-autoplot(pred.us)+
  ggtitle("United States")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

UK.plot<-autoplot(pred.uk)+
  ggtitle("United Kingdom")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

GE.plot<-autoplot(pred.ge)+
  ggtitle("Germany")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

CH.plot<-autoplot(pred.ch)+
  ggtitle("China")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

RU.plot<-autoplot(pred.ru)+
  ggtitle("Russia")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

FR.plot<-autoplot(pred.fr)+
  ggtitle("France")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

# Save the plots 
ggsave("US.png",US.plot,width = 10,height = 7.5)
ggsave("UK.png",UK.plot,width = 10,height = 7.5)
ggsave("GE.png",GE.plot,width = 10,height = 7.5)
ggsave("CH.png",CH.plot,width = 10,height = 7.5)
ggsave("RU.png",RU.plot,width = 10,height = 7.5)
ggsave("FR.png",FR.plot,width = 10,height = 7.5)

# view the plots
US.plot
UK.plot
GE.plot
CH.plot
RU.plot
FR.plot
```

Next, plot the predictions on the same scales, from 2.75e+10 to 6.5e+11, to compare the predicted CO2 emissions among the six countries. 
```{r plot the predictions based on the same scale (countries)}
# devtools::install_github('cttobin/ggthemr')
US.scaled.plot<-autoplot(pred.us)+
  ggtitle("United States")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(2.75e+10, 6.5e+11))

UK.scaled.plot<-autoplot(pred.uk)+
  ggtitle("United Kingdom")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(2.75e+10, 6.5e+11))

GE.scaled.plot<-autoplot(pred.ge)+
  ggtitle("Germany")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(2.75e+10, 6.5e+11))

CH.scaled.plot<-autoplot(pred.ch)+
  ggtitle("China")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(2.75e+10, 6.5e+11))

RU.scaled.plot<-autoplot(pred.ru)+
  ggtitle("Russia")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(2.75e+10, 6.5e+11))

FR.scaled.plot<-autoplot(pred.fr)+
  ggtitle("France")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(2.75e+10, 6.5e+11))

# Save the plots 
ggsave("US.scaled.png",US.scaled.plot,width = 10,height = 7.5)
ggsave("UK.scaled.png",UK.scaled.plot,width = 10,height = 7.5)
ggsave("GE.scaled.png",GE.scaled.plot,width = 10,height = 7.5)
ggsave("CH.scaled.png",CH.scaled.plot,width = 10,height = 7.5)
ggsave("RU.scaled.png",RU.scaled.plot,width = 10,height = 7.5)
ggsave("FR.scaled.png",FR.scaled.plot,width = 10,height = 7.5)

US.scaled.plot
UK.scaled.plot
GE.scaled.plot
CH.scaled.plot
RU.scaled.plot
FR.scaled.plot
```

### Predictions on continents
#### Time series analysis on five continents
```{r extract and prepare the continent data}
CO2$Continent <- countrycode(sourcevar = CO2[,"Country"],
                                  origin = "country.name",
                                  destination = "continent")
CO2_forecasting<-CO2 %>% 
  dplyr::select(Country, Year, CO2.emission..Tons.,Continent)
# make sure there is no NA
dim(CO2_forecasting)
dim(na.omit(CO2_forecasting))
# Use 1992-2020 data to forecast
CO2_no_year_min[which(CO2_no_year_min$Year==max(CO2_no_year_min[]$Year)),]
CO2_forecasting<-CO2_forecasting %>% 
  filter(Year>1991) %>% 
  group_by(Country)
# select continent
CO2_forecasting_continent<-CO2_forecasting %>% 
  dplyr::group_by(Year, Continent) %>%
  dplyr::summarise(Sum=sum(CO2.emission..Tons.))
CO2_forecasting_continent<-na.omit(CO2_forecasting_continent)
# View(CO2_forecasting_continent)
```

```{r time series model fits for five continents}
Asia<-CO2_forecasting_continent %>% 
  filter(Continent=="Asia")
Europe<-CO2_forecasting_continent %>% 
  filter(Continent=="Europe")
Africa<-CO2_forecasting_continent %>% 
  filter(Continent=="Africa")
Americas<-CO2_forecasting_continent %>% 
  filter(Continent=="Americas")
Oceania<-CO2_forecasting_continent %>% 
  filter(Continent=="Oceania")

# forecasting
# Asia
ts_data_as = ts(Asia$Sum,start=1992,frequency = 1)
arima.ts.as = auto.arima(ts_data_as)
pred.as = forecast(arima.ts.as,h=30)
# EU
ts_data_eu = ts(Europe$Sum,start=1992,frequency = 1)
arima.ts.eu = auto.arima(ts_data_eu)
pred.eu = forecast(arima.ts.eu,h=30)
# AF
ts_data_af = ts(Africa$Sum,start=1992,frequency = 1)
arima.ts.af = auto.arima(ts_data_af)
pred.af = forecast(arima.ts.af,h=30)
# AM
ts_data_am = ts(Americas$Sum,start=1992,frequency = 1)
arima.ts.am = auto.arima(ts_data_am)
pred.am = forecast(arima.ts.am,h=30)
# OC
ts_data_oc = ts(Oceania$Sum,start=1992,frequency = 1)
arima.ts.oc = auto.arima(ts_data_oc)
pred.oc = forecast(arima.ts.oc,h=30)
```

#### Plot the predicted CO2 emissions for the five continents
We expect to see what's the increasing or decreasing trend of the CO2 emissions in the next 30 years. 
```{r plot predictions on CO2 emissions by continents}
Asia.plot<-autoplot(pred.as)+
  ggtitle("Asia")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

Europe.plot<-autoplot(pred.eu)+
  ggtitle("Europe")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

Africa.plot<-autoplot(pred.af)+
  ggtitle("Africa")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

Americas.plot<-autoplot(pred.am)+
  ggtitle("Americas")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

Oceania.plot<-autoplot(pred.oc)+
  ggtitle("Oceania")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))

# save the plots
ggsave("Oceania.png",Oceania.plot,width = 10,height = 7.5)
ggsave("Asia.png",Asia.plot,width = 10,height = 7.5)
ggsave("Europe.png",Europe.plot,width = 10,height = 7.5)
ggsave("Africa.png",Africa.plot,width = 10,height = 7.5)
ggsave("Americas.png",Americas.plot,width = 10,height = 7.5)

Asia.plot
Europe.plot
Americas.plot
Oceania.plot
Africa.plot
```

Next, plot the predictions on the same scales, from 1e+10 to 2e+12, to compare the predicted CO2 emissions among the five continents. 
```{r plot the predictions based on the same scale (continents)}
Asia.scaled.plot<-autoplot(pred.as)+
  ggtitle("Asia")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(1e+10,2e+12))

Europe.scaled.plot<-autoplot(pred.eu)+
  ggtitle("Europe")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(1e+10,2e+12))

Africa.scaled.plot<-autoplot(pred.af)+
  ggtitle("Africa")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(1e+10,2e+12))

Americas.scaled.plot<-autoplot(pred.am)+
  ggtitle("Americas")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(1e+10,2e+12))

Oceania.scaled.plot<-autoplot(pred.oc)+
  ggtitle("Oceania")+
  labs(x = "Year", y = "CO2 Emission")+
  theme(axis.title.x = element_text(vjust = 0),
        axis.title.y = element_text(vjust = 2))+
  ylim(c(1e+10,2e+12))

# save the plots
ggsave("Oceania.scaled.png",Oceania.scaled.plot,width = 10,height = 7.5)
ggsave("Asia.scaled.png",Asia.scaled.plot,width = 10,height = 7.5)
ggsave("Europe.scaled.png",Europe.scaled.plot,width = 10,height = 7.5)
ggsave("Africa.scaled.png",Africa.scaled.plot,width = 10,height = 7.5)
ggsave("Americas.scaled.png",Americas.scaled.plot,width = 10,height = 7.5)

Asia.scaled.plot
Europe.scaled.plot
Americas.scaled.plot
Oceania.scaled.plot
Africa.scaled.plot
```
