---
title: "Final"
author: "Jennifer Liu, David Kook"
date: "2022-11-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plyr)
library(dplyr)
library(data.table)
library(countrycode)
```

Read in the raw data
```{r}
CO2 <- read.csv("CO2_emission_by_countries.csv", na = c("", "NA"), sep=",")
```

Check for all the NA entries
```{r}
rawCO2<-CO2 
rawCO2%>% summarise_all(~ sum(is.na(.)))
```

Find all the countires with NA in each column
```{r}
naCode <- rawCO2 %>% filter(is.na(Code)) %>% distinct(Country) 
naCall <- rawCO2 %>% filter(is.na(Calling.Code)) %>% distinct(Country)
naPopu <- rawCO2 %>% filter(is.na(Population.2022.)) %>% distinct(Country)
naArea <- rawCO2 %>% filter(is.na(Area)) %>% distinct(Country)
naXPre <- rawCO2 %>% filter(is.na(X..of.World)) %>% distinct(Country)
naDens <- rawCO2 %>% filter(is.na(Density.km2.)) %>% distinct(Country)

allNA <- bind_rows(naCode,naCall,naPopu,naArea,naXPre,naDens)
allNA <- distinct(allNA) %>% arrange(Country)
```

Combine all the table above to better see which values are missing
```{r}
nacheck<-function(str){
  df<- get(str)
  df$UniC <- df$Country
  colnames(df)[1] = str
  return(df)
}
naCode<- nacheck("naCode")
naCall<- nacheck("naCall")
naPopu<- nacheck("naPopu")
naArea<- nacheck("naArea")
naXPre<- nacheck("naXPre")
naDens<- nacheck("naDens")
```

join all the table together
```{r}
allNATable<-allNA %>% full_join(naCode,by = c("Country" = "UniC")) %>% 
  full_join(naCall,by = c("Country" = "UniC")) %>%
  full_join(naPopu,by = c("Country" = "UniC")) %>%
  full_join(naArea,by = c("Country" = "UniC")) %>%
  full_join(naXPre,by = c("Country" = "UniC")) %>%
  full_join(naDens,by = c("Country" = "UniC"))
view(allNATable)
```

David & Will
1.1 General overview of data, any visual or numerical differences.
Finding the first year in history that countries began emitting CO2
```{r}
#finding and removing NAs
CO2 <- na.omit(CO2)
View(CO2)

#filter emissions 0 and use tidyverse groupby country minimum year
CO2_no0<-CO2 %>% group_by(min(Year)) %>% filter(CO2.emission..Tons.!=0)

CO2_no0 %>% 
    group_by(Country) %>% 
    slice(which.min(Year))

CO2_no0 <- data.table(CO2_no0)

#removing duplicates country names
CO2_duplicates <- CO2_no0[!duplicated(CO2_no0$Country),]
CO2_duplicates 

#finding  firstyear where CO2 is nonzero value
CO2_no_year_min <- CO2_no0[, list(Year = min(Year)), by = Country] 
CO2_no_year_min[]

#graphic of start years for different countries
plot(CO2_no_year_min$Year,main="Earliest Year with CO2 Data",
     xlab="Index",ylab="Year")
abline(h=1950,col="blue")
```

Finding total CO2 emissions per country & CO2 emissions per country per person
```{r}
#finding CO2 emission/population
CO2total_byCountry <- aggregate(CO2_no0$CO2.emission..Tons., by=list(Category=CO2_no0$Country), FUN=sum) 

#renaming columns
colnames(CO2total_byCountry)[1] <- "Country"
colnames(CO2total_byCountry)[2] <- "CO2.Total"
CO2total_byCountry #checking that the column renaming worked

#joining columns together to conduct calculations
CO2total_byCountry <- CO2total_byCountry %>% inner_join(CO2_duplicates) %>% select("Country","CO2.Total","Population.2022.")

#dividing total CO2 emissions per country by the country's population
CO2total_byCountry$CO2perPop <- NULL
CO2total_byCountry$CO2perPop <- CO2total_byCountry$CO2.Total/CO2total_byCountry$Population.2022.
CO2total_byCountry

#renaming column names
colnames(CO2total_byCountry)[2] <- "CO2.Total.Per.Country"
colnames(CO2total_byCountry)[4] <- "CO2.per.Person.per.Country"

summary(CO2total_byCountry)
CO2total_byCountry 

```
 
Aim 1.2A: To visualize the cumulative CO2 emissions by country and continent
```{r,message=FALSE}
CO2total_byCountry$continent <- countrycode(sourcevar = CO2total_byCountry[,"Country"],
                                  origin = "country.name",
                                  destination = "continent")

#summing the total CO2 emissions by geographic region: Oceania, Africa, Asia, Americas, and Europe (Antartica was not included in the dataset)
CO2total_byContinent <- aggregate(CO2total_byCountry$CO2.Total.Per.Country, by=list(Category=CO2total_byCountry$continent), FUN=sum)
CO2total_byContinent 
colnames(CO2total_byContinent) <- c("Geographic_Region","Total_CO2_Emissions")
barplot(CO2total_byContinent$Total_CO2_Emissions~CO2total_byContinent$Geographic_Region)

library(tmap)
library(sf)
library(raster)
library(spData)
library(maptools)
library(rworldmap)
#loading the 
data("World")
data("wrld_simpl")
data("countryExData")
```

Create a map visualizing the amount of CO2 emission by country
```{r}
world <- map_data("world")
View(world)
#Add the data you want to map countries by to world
#In this example, I add lengths of country names plus some offset
```

Creating a barplot & world map to display CO2 emissions per country 
```{r}
which(CO2total_byCountry$Country=="United States")
CO2total_byCountry$Country[171] <- "USA" #renaming "United States" to "USA", to ensure uniformity in the naming 
which(world$region=="USA")
CO2total_byCountry[171,]

#renaming "region" in the data we are importing to "country"
world <- inner_join(world,CO2total_byCountry,by=c("region"="Country"))

#barplot of CO2 total emissions in world map
continent_barplot <- ggplot(data=subset(CO2total_byCountry,!is.na(continent)),aes(x=continent,y=CO2.Total.Per.Country,fill=continent)) + geom_bar(stat="identity",position="dodge")

continent_barplot <- continent_barplot + labs(x="Continent",y="Total CO2 Emissions (Tons)",title="CO2 Totals by Continent",fill="Continent")
continent_barplot

#graphically presenting CO2 emission in a world map
gg <- ggplot() + theme(legend.position="bottom",legend.key.size = unit(0.5, "cm"),legend.key.width = unit(1,"cm")) + labs(fill="CO2 Total(tons)/Country",x="Longitude",y="Latitude")
gg <- gg + geom_map(data=world, map=world, aes(map_id=region,x=long, y=lat, fill=CO2.Total.Per.Country))

#changing features of world map (color, axises scales, adding a title, etc)
gg <- gg + scale_fill_gradient(low = "green", high = "red", guide = "colourbar",aesthetics = "fill") + scale_y_continuous(limits = c(-100,100)) + scale_x_continuous(limits = c(-175,175)) + scale_colour_continuous(labels="scientific")
gg <- gg + coord_equal() + ggtitle("World View of CO2 Emissions")
gg

```

Analyze the top 3 countries that emit the most CO2 by geographic region
```{r}
#finding and plotting the top 3 countries in each geographic region that emit the most CO2
top10Countries <- CO2total_byCountry[order(-CO2total_byCountry$CO2.Total.Per.Country),] #get CO2 emissions per country in descending order

#finding the top 3 countries in each geographic region that emit the most CO2
top3Africa<-head(top10Countries[top10Countries$continent=="Africa",],3)
top3Americas<-head(top10Countries[top10Countries$continent=="Americas",],3)
top3Asia<-head(top10Countries[top10Countries$continent=="Asia",],3)
top3Europe<-head(top10Countries[top10Countries$continent=="Europe",],3)
top3Oceania<-head(top10Countries[top10Countries$continent=="Oceania",],3)

top3pContinent <- rbind(top3Africa,top3Americas,top3Asia,top3Europe,top3Oceania)

colnames(top3pContinent)

#soring the data into a variable for plotting
mostEmissions <- ggplot(top3pContinent,aes(x=continent,y=CO2.Total.Per.Country))
 
colnames(top10Countries)
library(ggrepel) 
library(magick)
library(scales)

colnames(top3pContinent)

#plotting the data
mostEmissions +
  geom_text_repel(aes(label = Country), size = 2.5)+
  geom_point(aes(colour=CO2.Total.Per.Country))+
  ggtitle(expression(atop("Top 3 CO2 Country Emissions per Continent",
                          atop(italic("CO2 Total from Available Data"),""))))+
  xlab("Continent") +
  ylab("CO2 Emissions in Tons")+
  theme(text = element_text(family = 'serif'))+
  theme(plot.title = element_text(hjust=0.5))

```

C02 Emissions by Landmass:
```{r}
cleanedC02 <- read.csv("~/Desktop/QBS/QBS181/FinalProject/QBSsquad/CO2_Cleaned_V2.csv",sep=",") #identified countries with NA values, then excluded the lower 50% of the countries who contributed the least CO2 emissions

View(cleanedC02)
colnames(cleanedC02)

#joining country, total area of each country, percent of global land mass that the country takes up, and the total CO2 emissions per country into one table
cleanedC02 <- cleanedC02[!duplicated(cleanedC02$Country),]
cleanedC02 <- inner_join(cleanedC02,CO2total_byCountry) %>% select("Country","Area","X..of.World","CO2.Total.Per.Country")

View(cleanedC02)

#removing the percent from each of the percent total landmass column
cleanedC02$X..of.World <- as.numeric(gsub("[\\%,]","",cleanedC02$X..of.World))
str(cleanedC02$CO2.Total.Per.Country)

#summary statistics of the relationship: country land mass vs. total CO2 emissions per country
summary(landmass_C02_model)

#plotting a scatter plot for country land mass vs. total CO2 emissions per, and finding the correlation coefficient (R^2 value), to assess the strength of the correlation
plot(cleanedC02$CO2.Total.Per.Country~cleanedC02$Area,xlab="Country_Area",ylab="CO2 Total in Tons")
title(main="CO2 Total vs. Country Area")
legend("topleft",legend=paste("R-squared is ",format(summary(landmass_C02_model)$r.squared,digits=3)))

#creating a linear regression of country land mass vs. total CO2 emissions per country and plotting the regression
mod1 <- lm(Area~CO2.Total.Per.Country,data=cleanedC02)
plot(mod1)
ggplot(data=cleanedC02,aes(x=Area,y=CO2.Total.Per.Country))+geom_point(stat="identity")
```

C02 Emissions in Developed vs. Non-developed:
```{r}
library(RCurl)
library(httr)
dat1 <- read.csv("~/Desktop/QBS/QBS181/FinalProject/CO2_emission_by_countries.csv", sep=",")
x <- getURL("https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv")
dat2 <- read.csv(text = x) #importing gdp data per country
class(dat2)

#converting SQL data into base R data
library(sqldf)
merged.dat <- sqldf("SELECT d1.Country, d1.year, d1.`CO2.emission..Tons.`,
d2.gdp, d2.co2_per_gdp, d2.co2_per_capita, d2.coal_co2, d2.coal_co2_per_capita
      FROM dat1 AS d1
      JOIN dat2 AS d2
      ON d1.Country=d2.country
      AND d1.Year=d2.year")
#View(merged.dat)
#colnames(merged.dat)


merged.dat
merged.dat <- na.omit(merged.dat)
```

Finding average CO2 emissions via coal burning per country & per capita:
```{r}
#finding average coal CO2 emissions per country
avgCoalC02 <- aggregate(merged.dat$coal_co2,by=list(Category=merged.dat$Country), FUN=mean)

#finding the average coal CO2 emissions per capital per country
avgCoalC02perCapita <- aggregate(merged.dat$coal_co2_per_capita,by=list(Category=merged.dat$Country), FUN=mean)
```

Finding gdp per capita per country:
```{r}
#finding the average CO2 emissions per country
avgCO2byCountry <- aggregate(merged.dat$CO2.emission..Tons.,by=list(Category=merged.dat$Country), FUN=mean)

#finding the gdp per country
avgGDP <- aggregate(merged.dat$gdp,by=list(Category=merged.dat$Country), FUN=mean)

#finding the average
avgC02perGDP <- aggregate(merged.dat$co2_per_gdp,by=list(Category=merged.dat$Country), FUN=mean)

#calculating the gdp per capita for each country
avgC02perCapita <- aggregate(merged.dat$co2_per_capita,by=list(Category=merged.dat$Country), FUN=mean)

#omitting na values and finding the top 10 CO2 emissions per capita per country 
avgC02perCapita <-na.omit(avgC02perCapita)
avgC02perCapita_10 <- avgC02perCapita[order(avgC02perCapita$x),]
avgC02perCapita_10 <- head(avgC02perCapita_10,10)

#plotting top 10 CO2 emissions per capita per country
capitaplot <- ggplot(avgC02perCapita_10,
                     aes(Category,x))+
  geom_bar(stat="identity",fill="lightblue") +
  theme(axis.text.x = element_text(size = 11)) +
  scale_x_discrete(guide = guide_axis(n.dodge = 3))+
  xlab("Country")+
  ylab("C02 per Capita")+
  ggtitle(expression(atop("C02 per Capita per Country",
                          atop(italic("Top 10 Countries"),"")))) +
  theme_gray()+
  theme(plot.title = element_text(hjust=0.5))+
  theme(text = element_text(family = 'serif'))
  
capitaplot
```

C02 by GDP:
```{r}
colnames(avgGDP) <- c("Country","GDP_Average") #renaming column names

#merging country, GDP_Average, and Population 2022 columns into one table
avgGDP <- avgGDP %>% inner_join(CO2_duplicates) %>% select("Country","GDP_Average","Population.2022.")
avgGDP <- avgGDP %>% mutate(GDPperCapita = GDP_Average/Population.2022.) 

#creating ordinal data for developed vs. developing
#changed developed threshold to $9000 per capita because original GDP threshold was too conservative
develop_ord <- ifelse(avgGDP$GDPperCapita>=9000,"developed",
                      ifelse(avgGDP$GDPperCapita<9000,"developing",0))
avgGDP$DevelopStatus <- develop_ord

#merging C02 Data
avgGDP <- avgGDP %>% inner_join(CO2_duplicates) %>% select("Country","GDP_Average","Population.2022.","GDPperCapita","CO2.emission..Tons.","DevelopStatus")

#categorizing developed nations into a developed variable
#categorizing developing nations into a developing variable
developed <- avgGDP[avgGDP$DevelopStatus=="developed",]
developing <- avgGDP[avgGDP$DevelopStatus=="developing",]
sumC02Developed <- sum(developed$CO2.emission..Tons.)
sumC02Developing <- sum(developing$CO2.emission..Tons.)

#creating a table and a barplot comparing developed vs developing nations total CO2 emissions
DevelopvsDeveloping <- matrix(ncol=2,nrow=1)
colnames(DevelopvsDeveloping)<-c("Developed","Developing")
DevelopvsDeveloping <- c(sumC02Developed,sumC02Developing)
DevelopvsDeveloping <- as.table(DevelopvsDeveloping)
DevelopvsDeveloping
barplot(DevelopvsDeveloping,names.arg = c("Developed","Developing"))


```


